# Exam 1 Practice

You should expect to be able to do 7 to 10 of these type problems for your exam (50 minutes).

1. Load the `rock` data set. How many observations and variables are in the `rock` data set?
2. What was the purpose of this dataset?
3. What are the units of measurement for `peri`?
4. In the `rock` data set what is the standard deviation of `area`? **`r sd(rock$area)`**
5. How many rocks have an `area` of greater than 4333? **`r length(rock$area[rock$area > 4333])`**
6. How many rocks have an `area` of greater than 4333 and less than 7333? **`r length(rock$area[rock$area > 4333 & rock$area < 7333])`**
7. Create a variable called `good.fun` which is area divided by perimeter squared.  What is the average of `good.fun`? **`r mean(rock$area/(rock$peri)^2)`**
8. Add a new observation that to this dataframe that has an `area` of 17777, `peri` of 7777, `shape` of 7777, and a `perm` of 7777.  Can you do this?
9. What is the mean of area for including the new rock you have added?
10. Using the `women` data set, what is the average BMI for all women?
11. What is the average BMI for `women`? 
12. What is the correlation between height and weight in the `women` dataset?
13. Make a histogram plot of the area variable from `rock`.
14. Use a bivariate plot of `rock` data to identify which variable are most correlated.  Confirm your answer with a correlation table.  Were you right?
15. Do this math in R: 3333 squared divided by the square root of 78.
16. Make this sequence using `rep` and `seq` functions: **1 2 3 1 2 3 1 2 3 1 2 3**
17. Make this sequence using `rep` and `seq` functions:  **`r rep(seq(97,89),3)`**
18. Make this plot:
```{r echo=FALSE}
plot(women$height, women$weight)
```
19. What is the heaviest car in the `mtcars` dataset in kg?
20. How many cars in `mtcars` dataset have more than 4 cylinders?
21. Create this dataframe called `Gerry`:
```{r echo=FALSE}
Gerry <- data.frame(x=20:29,y=120:129, z= 1120:1129)
```
22. Add column called `aa` to `Gerry` which are the numbers **`r 11120:11129`**
23. Have fun.  You are doing `awesome!`
