# dplyr

## Introduction

For more help **PLEASE** check out  [Introduction to dplyr](https://dplyr.tidyverse.org/articles/dplyr.html) introducing the key functionality of the dplyr package.

https://dplyr.tidyverse.org/articles/dplyr.html

> Your life is about to change.  For the better, even.


## A Neat Resource

* [RStudio's Data Wrangling Cheat Sheet](http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf) for dplyr and tidyr


## Single table verbs

`dplyr` aims to provide a function for each basic verb of data manipulation. These verbs can be organised into three categories based on the component of the dataset that they work with:

Rows:

- `filter()` chooses rows based on column values.
- `slice()` chooses rows based on location.
- `arrange()` changes the order of the rows.
    
Columns:

- `select()` changes whether or not a column is included.
- `rename()` changes the name of columns.
        `mutate()` changes the values of columns and creates new columns.
- `relocate()` changes the order of the columns.
    Groups of rows:
- `summarise()` collapses a group into a single row.  It’s not that useful until we learn the `group_by()` verb below.

## The pipe

All of the `dplyr` functions take a data frame (or tibble) as the first argument.  You can use the pipe to rewrite multiple operations that you can read left-to-right, top-to-bottom (reading the pipe operator as “then”).

> What is this: `%>%`?

## Loading `dplyr` 

```{r eval=FALSE}
# You should already have done this but you'll need it
install.packages("dplyr")
```


```{r echo=FALSE}
# load packages
suppressMessages(library(dplyr))
```

## `starwars` examples

```{r eval=TRUE}
library(dplyr)

starwars %>% 
  filter(species == "Droid")


starwars %>% 
  select(name, ends_with("color"))


starwars %>% 
  mutate(name, bmi = mass / ((height / 100)  ^ 2)) %>%
  select(name:mass, bmi)


starwars %>% 
  arrange(desc(mass))

starwars %>%
  group_by(species) %>%
  summarise(
    n = n(),
    mass = mean(mass, na.rm = TRUE)
  ) %>%
  filter(
    n > 1,
    mass > 50
  )
```

